---
title: "Rise or fall of programming languages"
author: "Jason"
date: "2022-11-01"
output: html_document
---
## Introduction 
This is my data analysis for stacked overflow data from 2008-2020. From this data one can deduce the most popular programming language used by developers based on the amount of questions asked. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reading the csv file. 
```{r}
StackofData <- read_csv("/Users/ice50l/Downloads/Rise and Fall of Programming Languages/datasets/stack_overflow_data.csv")
```

Inspecting the Dataset/Converting fractions to percentages. 
```{r}
print(StackofData)
r_percentage <- StackofData %>%
  mutate(fraction = number / year_total*100)
r_percentage
```

Viewing the data 
```{r}
View(StackofData)
```

Find the top 5 tags that were searched/asked from 2015-2020
```{r}
sorted_tags <- StackofData %>%
  filter(year >= 2015) %>% 
  group_by(tag) %>% 
  summarize(tag_total = sum(number)) %>% 
  arrange(desc(tag_total))
sorted_tags
highest_tags <- head(sorted_tags$tag, n = 5)
highest_tags
```

Sorting the lowest searched/asked from 2015-2020
```{r}
sorted_tagsLOW <- StackofData %>%
  filter(year >= 2015) %>% 
  group_by(tag) %>% 
  summarize(tag_total = sum(number)) %>% 
  arrange((tag_total))
sorted_tagsLOW

LowestTags<- head(sorted_tagsLOW$tag, n = 5)
LowestTags
```


Filtering highest/lowest searched/asked from 2015-2020
```{r}
by_tag_subset <- r_percentage %>% 
  filter(tag %in% highest_tags, year >= 2015)

LowTag <- r_percentage %>%
  filter(tag %in% LowestTags, year >= 2015 )
```


Plotting tags 2015-2020 for highest tags. 
```{r}
# Plot tags over time on a line plot using color to represent tag
ggplot(by_tag_subset, aes(x = year, y = fraction, color = tag)) +
  geom_line() + labs(y = "Perecent of Total Question Asked")
```

Plotting percent of total questions asked from the total amount for selected programming languages (r, scala, python, java, sql)
```{r}
Sample <- c("r", "scala", "python", "java", "sql")
SampleOT <- r_percentage %>%
  filter(year >= 2015) %>%
  filter(tag %in% Sample)

#Plotting the sample over time.
ggplot(SampleOT, aes(x = year, y = fraction, color = tag)) + geom_line() + labs(y = 'Percent of Total Questions Asked')
```

Plotting Android vs IOS questsions asked/searched from 2010-2020

```{r}
IDroid <- c('ios', 'android')
tagIDroid <- r_percentage %>%
  filter(tag %in% IDroid,year >= 2010)

ggplot(tagIDroid, aes(x = year, y = fraction, color = tag)) + geom_line() + labs(y = 'Percent of Total Questions Asked')
```   

## Conclusions

We can deduce that over the years 2015-2020, python has the most rapid growth out of every other programming language. While Javascript remaioned about costant over 5 years. Some langauges/OS such as Android has since fallen in question popularity in the 5 years. 

1. We can conclude taht Python has the biggest growth out of any other tags.
2. Java has fallen in popularity in 5 years, while scala, r, and sql over 2015-2020 make up approximately 1% of total questions asked on StackOverflow. 
3. IOS has been lagging behind Android in questions asked over 2010-2020, IOS never has exceeded 2% of the total compared to Android, which has been hovering over 2% for many years. 























